Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-22 paper/icml_drau/advup.tex
.

Contrary to the action state value function,
this
rescaled advantage function converges when $\deltat\to 0$
to an action-dependent quantity which, if it can be learned, contains the
necessary information for policy improvement \TODO{See Thm... in the appendix; we
should put the theorem in the appendix, because otherwise the theorem is
just a paraphrase of the previous sentence}.

@@

\begin{theorem}
	\TODO{We need a bit more details: we probably want to assume that states and actions are
	drawn from a single fixed trajectory, and that we are learning for a fixed, constant policy.}
	Let $\eta^V_\deltat = \alpha^V \deltat^\beta$ (resp. $\eta^A \deltat^\beta$) be the learning rate
	of $V$ (resp. $A$), when training on a single exploratory trajectory. Then
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters grow arbitrarily large in a single timestep as $\deltat$
			goes to $0$. \TODO{not quite true, in a single
			step the update is bounded. But the parameters
			become arbitrary large in an arbitrarily small
			physical time.}
		\item If $\beta > 1$, parameter trajectories become
		stationary as
			$\deltat$ goes to $0$.
	\end{itemize}
	\label{th:cont-params}
\end{theorem}

@@

>>>>

Contrary to the action state value function,
this
rescaled advantage function converges when $\deltat\to 0$
to an action-dependent quantity which, if it can be learned, contains the
necessary information for policy improvement. A formal statement is given in the supplementary material.

@@

\begin{theorem}
	\TODO{Assume a continuous action space. Let $(s_t, a_t)_{t\geq 0}$ be a
	time continuous exploratory trajectory. Let $\pi$ be a sufficiently
regular policy, and $\bar{A}$ and $\bar{V}$ be regular function approximators.}
	Let $\eta^V_\deltat = \alpha^V \deltat^\beta$ (resp. $\eta^A = \alpha^A \deltat^\beta$) be the learning rate
	of $V$ (resp. $\bar{A}$). Then, when learning on $(s_t, a_t)_{t\geq 0}$, 
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters grow arbitrarily large in after an arbitrarily small physical time when $\deltat$ goes to $0$. \TODO{not quite true, in a single
			step the update is bounded. But the parameters
			become arbitrary large in an arbitrarily small
			physical time.}
		\item If $\beta > 1$, parameter trajectories become
		stationary as
			$\deltat$ goes to $0$.
	\end{itemize}
	\label{th:cont-params}
\end{theorem}

@@

%%% Local Variables:
%%% TeX-master: "icml_drau"
%%% End:

### 2019-01-22 paper/icml_drau/experiments.tex
.

In all settings, we use the learning algorithm described in
Alg.~\ref{alg:dau} and \TODO{supplementary, algo 1}. The variants of DDPG and DQN
used are described in the supplementary material, as well as all the hyperparameters
used. To make comparison fair, we use the properly rescaled discount
factor $\gamma_\deltat$ and reward $r_\deltat$ from \eqref{eq:def-gamma},
as well as scaled learning rates, for DDPG and DQN. \TODO{NDY not
convincing: precisely because DDPG and DQN are not invariant, they might
benefit from a different scaling.}

>>>>

In all settings, we use the learning algorithm described in
Alg.~\ref{alg:dau} and supplementary, Algo 1. The variants of DDPG and DQN
used are described in the supplementary material, as well as all the hyperparameters
used. To make comparison fair, we tested to different scalings for DDPG and DQN.
One where we use the properly rescaled discount
factor $\gamma_\deltat$ and reward $r_\deltat$ from \eqref{eq:def-gamma},
as well as scaled learning rates. One where we only scaled the discount factor
(to avoid shortsightedness), but left all other hyperaparameters constant
across time discretizations. Experiments on the latter yielded better result
and is presented hereby. Experiments on the first variant are presented in the
supplementary material. \TODO{NDY not
convincing: precisely because DDPG and DQN are not invariant, they might
benefit from a different scaling.}

### 2019-01-22 paper/icml_drau/framework.tex
.

A necessary condition for robustness of an algorithm for
near-continuous time MDPs is 
to remain viable when $\deltat \rightarrow 0$. Thus we will try to
make sure the various quantities involved converge to meaningful
limits when $\deltat\to 0$.

@@

When the timestep tends to $0$, one converges to the other.
  \begin{theorem}
    \label{th:conv-value}
   Under suitable assumptions \TODO{ref appendix}, for all $s \in {\cal
   S}$, one has
    %\begin{equation}
      %\label{eq:conv-value}
      $V^\pi_\deltat(s) = V^\pi(s) + \bigO(\deltat)$
    %\end{equation}
    when $\deltat\to 0$.
  \end{theorem}

\subsection{There is no $Q$-function in continuous time}

>>>>

A necessary condition for robustness of an algorithm for
near-continuous time MDPs is 
to remain viable when $\deltat \rightarrow 0$. Thus we will try to
make sure the various quantities involved converge to meaningful
limits when $\deltat\to 0$. \TODO{Formal statements are made in the following section. Proofs and technical assumptions can be found in the supplementary material.}

@@

When the timestep tends to $0$, one converges to the other.
  \begin{theorem}
    \label{th:conv-value}
    Assume that $r: {\cal S} \times {\cal A} \rightarrow \bb R$ is bounded, and
    that $s \rightarrow r(s, \pi(s))$ is lipschitz continuous, then
    for all $s \in {\cal S}$, one has
    %\begin{equation}
    %\label{eq:conv-value}
    $V^\pi_\deltat(s) = V^\pi(s) + \smallo(\deltat)$
    %\end{equation}
    when $\deltat\to 0$.
  \end{theorem}

\subsection{There is No $Q$-Function in Continuous Time}

### 2019-01-22 paper/icml_drau/supplementary.tex
.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{xcolor}

@@

\begin{document}
\section{Proofs}
We hereby give proofs for all the results presented in the paper.
The first result presented is a proof of convergence for discretized
trajectories.
\begin{theorem}
	Let $F: {\cal S} \times {\cal A} \rightarrow \bb{R}^n$ and $\pi: {\cal S}
	\rightarrow {\cal A}$ be the dynamic and policy functions. Assume that,
	for any $a$, $s \rightarrow F(s, a)$ and $s \rightarrow F(s, \pi(s))$
	are ${\cal C}^1$, bounded and $K$-lipschitz.  For
	a given $s_0$, define the trajectory $(s_t)_{t\geq 0}$ as the unique
	solution of the differential equation
	\begin{equation}
		\frac{ds_t}{dt} = F(s_t, \pi(s_t)).
		\label{eq:diff}
	\end{equation}
	For any $\deltat > 0$, define the discretized trajectory
	$(s_\deltat^k)_k$ recurrently as $s_\deltat^0 = s_0$,
	$s_\deltat^{k + 1}$ is the value at time $\deltat$ of
	the unique solution of
	\begin{equation}
		\frac{d\tilde{s}_t}{dt} = F(\tilde{s}_t, \pi(s_\deltat^k))
	\end{equation}
	with initial point $s_\deltat^k$.
	Then, there exists $C > 0$ such that, for every $t \geq 0$
	\begin{equation}
		\|s_t - s_\deltat^{\lfloor \frac{t}{\deltat} \rfloor}\|
		\leq \deltat \frac{C}{K}e^{Kt}.
	\end{equation}
	Notably, discretized trajectories converge pointwise to continuous trajectories.
	\label{th:traj-conv}
\end{theorem}
\begin{proof}
	\input{proof_traj.tex}
\end{proof}
In what follows, we will assume that the reward function $r: {\cal S} \times {\cal A} \rightarrow \bb{R}$
is bounded, to ensure existence of $V^\pi$ and $V^\pi_\deltat$ for all $\deltat$.\begin{theorem}
	Under suitable assumptions \TODO{ref appendix}, for all $s \in {\cal
	S}$, one has
	%\begin{equation}
	%\label{eq:conv-value}
	$V^\pi_\deltat(s) = V^\pi(s) + \bigO(\deltat)$
	%\end{equation}
	when $\deltat\to 0$.
	\label{th:conv-value}
\end{theorem}
\begin{proof}
	\input{proof_conv_value.tex}
\end{proof}

@@

	If for all $s$, $A^\pi(s, \pi'(s)) \geq 0$, then for all $s$, $V^\pi(s) \leq V^{\pi'}(s)$.
\end{theorem}
\begin{proof}
	\input{policy_improvement.tex}
\end{proof}
\begin{theorem}
	Discrete parameter trajectories converge to continuous trajectories.
\end{theorem}
\begin{proof}
	\input{proof_conv_params.tex}
\end{proof}
\section{Implementation details}
\end{document}

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algorithmic}

@@

\begin{document}
\section{Proofs}
We hereby give proofs for all the results presented in the paper.
The first result presented is a proof of convergence for discretized
trajectories.
\begin{theorem}
	Let $F: {\cal S} \times {\cal A} \rightarrow \bb{R}^n$ and $\pi: {\cal S}
	\rightarrow {\cal A}$ be the dynamic and policy functions. Assume that,
	for any $a$, $s \rightarrow F(s, a)$ and $s \rightarrow F(s, \pi(s))$
	are ${\cal C}^1$, bounded and $K$-lipschitz.  For
	a given $s_0$, define the trajectory $(s_t)_{t\geq 0}$ as the unique
	solution of the differential equation
	\begin{equation}
		\frac{ds_t}{dt} = F(s_t, \pi(s_t)).
		\label{eq:diff}
	\end{equation}
	For any $\deltat > 0$, define the discretized trajectory
	$(s_\deltat^k)_k$ recurrently as $s_\deltat^0 = s_0$,
	$s_\deltat^{k + 1}$ is the value at time $\deltat$ of
	the unique solution of
	\begin{equation}
		\frac{d\tilde{s}_t}{dt} = F(\tilde{s}_t, \pi(s_\deltat^k))
	\end{equation}
	with initial point $s_\deltat^k$.
	Then, there exists $C > 0$ such that, for every $t \geq 0$
	\begin{equation}
		\|s_t - s_\deltat^{\lfloor \frac{t}{\deltat} \rfloor}\|
		\leq \deltat \frac{C}{K}e^{Kt}.
	\end{equation}
	Notably, discretized trajectories converge pointwise to continuous trajectories.
	\label{th:traj-conv}
\end{theorem}
\begin{proof}
	\input{proof_traj.tex}
\end{proof}
In what follows, we will assume that the reward function $r: {\cal S} \times {\cal A} \rightarrow \bb{R}$
is bounded, to ensure existence of $V^\pi$ and $V^\pi_\deltat$ for all $\deltat$
.\begin{theorem}
	Assume that $r: {\cal S} \times {\cal A} \rightarrow \bb R$ is bounded, and
	that $s \rightarrow r(s, \pi(s))$ is $L_r$-lipschitz continuous, then
	for all $s \in {\cal S}$, one has
	%\begin{equation}
	%\label{eq:conv-value}
	$V^\pi_\deltat(s) = V^\pi(s) + \smallo(\deltat)$
	%\end{equation}
	when $\deltat\to 0$.
	\label{th:conv-value}
\end{theorem}
\begin{proof}
	\input{proof_conv_value.tex}
\end{proof}

@@

	If for all $s$, $A^\pi(s, \pi'(s)) \geq 0$, then for all $s$, $V^\pi(s) \leq V^{\pi'}(s)$.
\end{theorem}
\begin{proof}
	\input{policy_improvement.tex}
\end{proof}
\begin{theorem}
	Discrete parameter trajectories converge to continuous trajectories.
\end{theorem}
\begin{proof}
	\input{proof_conv_params.tex}
\end{proof}
\section{Implementation details}
\input{impl_details.tex}
\end{document}
